---
title: "Untitled"
author: "Marissa Dyck"
date: "2024-05-08"
output: html_document
---



# Autocorellation

```{r}
# Couldn't get this to work in purrr yet so using a loop to subset the data, create the plots, and save them all in one section... NEAT

buffer_frames <- list()

for (i in unique(covariates_grouped$buff_dist)){
  
  print(i)
  
  #Subset data based on radius
  df<-covariates_grouped %>%
    filter(buff_dist == i)
  
  #rename dataframe on the fly
  assign(paste("df", i, sep ="_"), df)
  
  #list of dataframes
  buffer_frames <-c (buffer_frames, list(df))
  
  #Subset data based on radius
  df<-covariates_grouped %>%
    filter(buff_dist == i) %>%
    select(where(is.numeric))
  
 #compute a correlation matrix (watch for errors)
 matrix <- cor(df)
 
 # print and save the correlation plot on the go
 # renaming for each buffer as we do
 png(file.path("figures/", paste("correlation_", i, ".png")))
 corrplot::corrplot(matrix,
                    type = 'upper',
                    tl.col = 'black',
                    title = paste0('Variable correlation plot at ', i))
 dev.off()
  
}
```

## Correlation plots

Now we need to make correlation plots for each buffer width to see what variables are correlated at a given spatial scale. We can use `purrr::map()` with the `chart.Correlation()` function from the *PerformanceAnalytics* package to make correlation plots with a specified method (e.g., pearson, spearman, etc.) That also show histograms and scatterplots of each variable.

```{r correlation plots pearson, warning=FALSE}

correlation_plots <- buffer_frames %>% 
  
  purrr::map(
    ~.x %>% 
      
      # select numeric variables only since we can't compute a r2 for non-numeric
      select_if(is.numeric) %>% 
      
      # use chart.correlation 
      chart.Correlation(.,
                        histogram = TRUE, 
                        method = "pearson")
  )
```

Although we are able to plot all of the histograms at once, with so many buffers and without an easy way to add titles to all of these plots in purrr it's hard to keep track of everything. So instead, I;ve added a section for each buffer width, with the correlation plot and outlining the variables that are autocorrelated and thus should not be included in the same model, it includes the r2 as well

## 250m

```{r cor 250, warning=FALSE}

buffer_frames$`250 meter buffer` %>% 
  
  select_if(is.numeric) %>% 
  
   # use chart.correlation 
      chart.Correlation(.,
                        histogram = TRUE, 
                        method = "pearson")

mtext('250 meter buffer', side = 3, line = 3)
  
```

* pipeline & transmission_lines 0.53    
* roads & veg_edges 0.71
* roads & lc_developed 0.57

## 500m

```{r cor 500, warning=FALSE}

buffer_frames$`500 meter buffer` %>% 
  
  select_if(is.numeric) %>% 
  
   # use chart.correlation 
      chart.Correlation(.,
                        histogram = TRUE, 
                        method = "pearson")

mtext('500 meter buffer', side = 3, line = 3)
  
```

* harvest &  lc_mixed 
* pipeline & transmission_lines 0.64  
* pipeline & roads 0.59   
* pipeline & veg edges 0.57   
* pipeline & wells 0.75   
* pipeline & lc_grassland 0.62    
* roads & veg edges 0.79    
* roads & wells 0.75    
* roads & lc_developed 0.75      
* roads & osm_industrial 0.72   
* veg edges & wells 0.61    
* veg edges & lc_developed 0.88      
* wells & lc_grassland 0.61       
* wells & lc_developed 0.67      
* coniferous & broadleaf -0.77    


## 750m

```{r cor 750, warning=FALSE}

buffer_frames$`750 meter buffer` %>% 
  
  select_if(is.numeric) %>% 
  
   # use chart.correlation 
      chart.Correlation(.,
                        histogram = TRUE, 
                        method = "pearson")

mtext('750 meter buffer', side = 3, line = 3)
  
```

* pipeline & transmission_lines 0.70       
* roads & veg_edges 0.74    
* roads & lc_developed 0.64    
* transmission_lines & lc_developed 0.51    







Now let's do the same thing with the landcover variables

```{r lc histograms, eval=FALSE}

lc_histograms <- buffer_frames %>% 
  
  purrr::imap(
    ~.x %>% 
      
      # filter to just the landcover variables 
      select(where(is.numeric) &
          starts_with('lc_class')) %>% 
      
      # pipe into hist.data.frame function to make histograms for each variable
      hist.data.frame(mtitl = paste0('Histograms of landcover variables at ', .y)))
```


Let's remove the 250 buffer and see what happens. To do this we can use the `discard_at()` function in the *purr* package which will remove a list element based on it's name (e.g. '250 meter buffer').

After that we will rerun the model selection again and see if it looks better
```{r remove 250 buffer}

black_bear_mods_no250 <- black_bear_mods  %>% 
  
  # purrr::discard_at will remove an item from a list
  purrr::discard_at('250 meter buffer')


# run model selection again
model.sel(black_bear_mods_no250)
```

> this looks much more realistic; the 500 m buffer is top model for black bears

So what we will do for each species is remove the 250 meter buffer for now since there are some data missing, and compare just the other buffer sizes that contain the full data set


### Model summary 500m

Let's take a look at the model summary for the top model

```{r black bear 500 summary}

summary(black_bear_mods_no250$`500 meter buffer`)
```


### Odds ratio plots

What we need to do now is extract the coefficient estimates from each top model, as well as the 95% CI so we can plot them for easier visualization and understanding of the data


#### Extract odds ratios

The `confint()` function will extract the coefficients and CI intervals from a model, so what we need to do is make a list of all the models, then use the `map()` function in *purrr* to apply the `confint()` function to all the models and get the coefficients. We want this to result in a tibble that has a column for the HFI feature (we aren't plotting the lc_class data for this report), the upper and lower CI, and the coefficient estimate.

In order to do this we have to do a bit of data wrangling, currently this isn't the most pleasing way to accomplish the desired outcome, but it works. 

```{r extract odds}

# This is also a dog shit way to do this but I need to get this done

# make a list of the top models for each species
top_models <- list(black_bear_mods$`250 meter buffer`,
                   caribou_mods$`3000 meter buffer`,
                   coyote_mods$`2250 meter buffer`,
                   fisher_mods$`1000 meter buffer`,
                   wolf_mods$`500 meter buffer`,
                   lynx_mods$`1250 meter buffer`,
                   moose_mods$`250 meter buffer`,
                   fox_mods$`4750 meter buffer`,
                   deer_mods$`1500 meter buffer`) %>%  
  
  # pipe into purrr to create coefficient table for all models
  purrr::map(
    ~.x %>% 
      
      # extract the coefficients and upper and lower CI
      confint() %>% 
      
      # format resulting object as a tibble data frame
      as_tibble() %>%
      
      # subset to just the HFI variables for these plots
      slice_head(n = 11) %>% 
      
      # remove first row which is the intercept
      slice_tail(n = 10) %>% 
      
      # add a column where we can put the feature names
      rowid_to_column() %>% 
      
      # rename the columns for plotting
      rename('lower' = `2.5 %`,
             'upper' = `97.5 %`,
             'estimate' = Estimate,
             'feature' = rowid) %>% 
      
      # rename the entries to features, need to look at the order the features are in from the model summary and ensure it matches
      mutate(feature = as.factor(feature),
             feature = recode(feature,
                              '1' = 'harvest',
                              '2' = 'pipeline',
                              '3' = 'roads',
                              '4' = 'seismic_lines',
                              '5' = 'seismic_lines_3D',
                              '6' = 'trails',
                              '7' = 'transmission_lines',
                              '8' = 'veg_edges',
                              '9' = 'wells',
                              '10' = 'osm_industrial'))) %>% 
  
  # set the names of each resulting tibble data frame to the species name
  purrr::set_names('Black bear',
                   'Caribou',
                   'Coyote',
                   'Fisher',
                   'Grey wolf',
                   'Lynx',
                   'Moose',
                   'Red fox',
                   'White-tailed deer')
```

#### Merge data frames

Now we have a data frame with the odds ratios for each species, but if we want these on a plot together we need them all in one data frame. 

To merge data into one data frame we can use the `list_rbind()` function from the *purrr* package which will take each element of the list and stack them on top of one another just like rbind does with data frames, and if we use the names_to argument we can extract the names of the list elements and assign them to a column so we know which data comes from which species model (list element)

In this code I also add a new column called uuid which contains the image id (uuid) for a phylopic silhouette of each species that I may want to use for plotting

[Phylopic.or](https://www.phylopic.org/) is an open source online database of silhouettes various contributors have created for use. There is an R package that works with this data called *rphylopic*; we can use the `get_uuid()` function from this package to extract the data for a silhouette for each species we want, which is what I've done here.


```{r rbind odds}

# combine all list elements 
coeffs_df_all <- list_rbind(top_models,
                            names_to = 'species') %>% 
  
  # change species to a factor for plotting
  mutate(species = as.factor(species),
         
         # add phylopic uuid for each species for plotting 
         # the uuid is extracted using getuuid with the species name as name = ''
         uuid = case_when(species == 'Black bear' ~ get_uuid(name = 'Ursus americanus'),
                          species == 'Caribou' ~ get_uuid(name = 'Rangifer tarandus'),
                          species == 'Coyote' ~ get_uuid(name = 'Canis latrans'),
                          species == 'Fisher' ~ get_uuid(name = 'Pekania pennanti'),
                          species == 'Grey wolf' ~ get_uuid(name = 'Canis lupus'),
                          species == 'Lynx' ~ get_uuid(name = 'Lynx lynx'),
                          species == 'Moose' ~ get_uuid(name = 'Alces alces'),
                          species == 'Red fox' ~ get_uuid(name = 'Vulpes vulpes'),
                          species == 'White-taield deer' ~ get_uuid(name ='Odocoileus virginianus'))) 
  
```

#### Plot odds all species

Now let's explore some different options to plot the coefficients

##### geom_phylopic

Let's try plotting all the species on one plot using `ggplot()`
```{r plot odds all}

# provide data and mapping aesthetics
ggplot(coeffs_df_all, aes(x = feature, 
                          y = estimate, 
                          group = uuid)) +
  
   geom_errorbar(aes(ymin = lower, 
                     ymax = upper, 
                     color = feature),
                width = 0.4,
                linewidth = 0.5,
                position = position_dodge(width = 0.9)) +
  
    # add points for each estimate for each covariate and use position = position_dodge to shift the points so all the species don't plot on top of one another
  geom_phylopic(aes(x = feature, 
                 y = estimate,
                 uuid = uuid),
             position = position_dodge(width = 0.9),
             size = 0.3)
  
  


```

This works but unfortunately is very messy with the images as the points since they don't all have the same centerline and are different sizes.

I try with the code below to select better silhouettes

```{r select silhouettes}

# combine all list elements 
coeffs_df_all <- list_rbind(top_models,
                            names_to = 'species') %>% 
  
  # change species to a factor for plotting
  mutate(species = as.factor(species),
         
         # add phylopic uuid for each species for plotting 
         # the uuid is extracted using getuuid with the species name as name = ''
         uuid = case_when(species == 'Black bear' ~ get_uuid(name = 'Ursus americanus'),
                           species == 'Caribou' ~ 'e6e864fd-8e3d-435f-9db3-dc6869c589f1', # caribou 5
                          species == 'Coyote' ~ 'e6a2fa4b-85df-43b4-989c-34a65ba7eee3', # coyote 6
                          species == 'Fisher' ~ '735066c6-2f3e-4f97-acb1-06f55ae075c9',
                          species == 'Grey wolf' ~ '76c8fdec-d0af-47b9-b949-dc610419d832', # wolf 5
                          species == 'Lynx' ~ '74eab34a-498c-4614-aece-f02361874f79', # moose 3
                          species == 'Moose' ~ '74eab34a-498c-4614-aece-f02361874f79',
                          species == 'Red fox' ~ 'd67d3bf6-3509-4ab6-819a-cd409985347e', # red fox 4
                          species == 'White-taield deer' ~ '6038e80c-398d-47b2-9a69-2b9edf436f64')) # white-tailed deer 6
```


Try plotting all again

```{r plot all odds 2}

ggplot(coeffs_df_all, aes(x = feature, 
                          y = estimate, 
                          group = uuid)) +
  
   geom_errorbar(aes(ymin = lower, 
                     ymax = upper, 
                     color = feature),
                width = 0.4,
                linewidth = 0.5,
                position = position_dodge(width = 1.2)) +
  
    # add points for each estimate for each covariate and use position = position_dodge to shift the points so all the species don't plot on top of one another
  geom_phylopic(aes(x = feature, 
                 y = estimate,
                 uuid = uuid),
             position = position_dodge(width = 1.2),
             size = 0.2)
  
  


```

Unfortunately they still don't center on the lines like they should so will try a different plot

##### Facet_wrap

```{r plot all odds facet_wrap}

ggplot(coeffs_df_all,
       aes(x = feature,
           y = estimate)) +
  
  # add points
  geom_point() +
  
  # add errorbar
  geom_errorbar(aes(ymin = lower, 
                     ymax = upper),
                width = 0.4,
                linewidth = 0.5) +
  
  facet_wrap(vars(species)) +
  
  # wrap long axis names
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  
  # adjust theme elements
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1),
        axis.ticks.x = element_blank())
```

##### Plot_model

If all else fails can use [plot_model function](https://cran.r-project.org/web/packages/sjPlot/vignettes/plot_model_estimates.html)

###### Example with black bear model

```{r plot_model}

library(sjPlot)

plot_model(black_bear_mods$`250 meter buffer`,
           vline.color = 'black')
```

